# Introduction {#intro}

```{r include = FALSE}
# Options
knitr::opts_chunk$set(echo = FALSE)

# Packages
library(tidyverse)
library(ISLR)

# Get raw data from the R package
data(Wage, Smarket, NCI60, package = "ISLR")
```

A brief definition of statistical learning: "Broadly speaking, supervised statistical learning involves building a statistical model for predicting, or estimating, an *output* based on one or more *inputs*" (emphasis original). 

Statistical Learning methods fall into two broad categories

* Supervised learning methods use data to model a known output based on related inputs
* Unsupervised learning methods find relationships between variables and structure in the data

## Datasets 

This chapter gives an overview of the data sets used in the book.

### `Wage`

Used for regression problem examples such as predicting wage based on age and education

```{r echo = TRUE}
glimpse(Wage)
```

### `Smarket`

Used for classification problem examples such as predicting whether a stock index will increase or decrease on any given day

```{r echo = TRUE}
glimpse(Smarket)
```


### `NCI60` 

Used for examples of clustering problems such as identifying related groups of cancer cells

```{r echo = TRUE}
str(NCI60)
```


```{r include = FALSE}
# Output csv files locally

if ( !file.exists("data/raw/Wage.csv") ) {
    write_csv(Wage, "data/raw/Wage.csv")
} 

if ( !file.exists("data/raw/Smarket.csv") ) {
    write_csv(Smarket, "data/raw/Smarket.csv")
} 

# if ( !file.exists("data/raw/NCI60.csv") ) {
#     write_csv(NCI60, "data/raw/NCI60.csv")
# } 
```
